### 2019-09-09 assignment_one/article/main.tex
Working on data generation.

\documentclass[article, a4paper, oneside, 12pt]{memoir}
\linespread{1.25}
\setlength{\parskip}{1em}
\setlength{\parindent}{0em}

\usepackage[]{mathtools}
\usepackage[]{amssymb}
\usepackage[]{libertine}
\usepackage[]{libertinust1math}
\usepackage[T1]{fontenc}
\usepackage[]{microtype}
\usepackage[]{geometry}
\usepackage[]{amsthm}
\usepackage[]{thmtools}
\usepackage[usedvipsnames]{xcolor}
\usepackage[]{bm}
\usepackage[noabbrev, capitalize]{cleveref}
\usepackage[]{amsmath}
\usepackage[]{commath}

\declaretheorem[style=remark, name=Remark,
mdframed={
}]{remark}

\title{\textsc{Regression Techniques \\
  FYS-STK4155 \\
Assignment One}}
\author{Ivar Haugal{\o}kken Stangeby}

\newcommand{\x}{\bm{x}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\X}{\bm{X}}

\newcommand{\mat}[1]{\bm{#1}}
\newcommand{\y}{\bm{y}}
\newcommand{\data}{\bm{\Omega}}
\newcommand{\cost}{\mathcal{C}}
\newcommand{\N}{\mathcal{N}}
\newcommand{\basis}{\mathcal{B}}
\newcommand{\set}[1]{\left\{\, #1 \right\}}

\begin{document}
  \maketitle

  \chapter{Introduction}

  Regression analysis is the act of estimating relationships among variables.
  In this project, we study various regression methods in more detail. In
  particular, we compare the \emph{ordinary least squares} (OLS) method with
  the \emph{Ridge regression} and \emph{Lasso regression} techniques. As we
  shall see, these three methods are all variations over the same theme. We
  start by testing the methods on noisy data sampled from a function known as
  \emph{Franke's} bivariate test function, which has been in widespread use in
  the evaluation of interpolation and data fitting techniques. Finally, we run
  regression on real terrain data, comparing the aforementioned methods.

  \chapter{Least Squares}

  We start by giving an introduction to ordinary least square regression. This
  method is perhaps the most widely used regression technique due to its
  simplicity, both in its derivation, and its implementation. The goal of
  regression analysis is to fit a \emph{model function} \( f(\x, \beta) \) to a
  set of \( n \) data points \( \data = (\x_i, y_i)_{i=1}^n \). A simple example
  is that of a linear polynomial with two parameters:
  \begin{equation}
    f(x, \beta) = \beta_0 + \beta_1 x.
  \end{equation}
  The \emph{model parameters} \( \beta \) are determined in order to minimize a
  suitable \emph{cost function} \(\cost(\data, \beta)\) which measures to which
  extent the model function manages to capture trends in the data \( \data \).

  One common cost function is the one which involves the sum of squared
  residuals (or squared errors):
  \begin{equation}
    \label{eq:mse}
    \cost(\data, \beta) = \frac{1}{n}\sum_{i = 1}^n \varepsilon_i^2 \coloneqq \frac{1}{n}\sum_{i=1}^n (y_i - \hat{y}_i)^2
  \end{equation}
  where \( \hat{y}_i \coloneqq f(\x_i, \beta) \). The method involving the
  minimization of this specific cost function is known as \emph{least squares}.

  \begin{remark}
    It is often assumed a priori that the data is infact generated from a noisy model, such that each \( y_i \) can be described 
    as
    \begin{equation}
      y_i = f(\x_i) + \varepsilon_i
    \end{equation}
    where each \( \varepsilon_i \sim \N(0, \sigma^2) \) is normally distributed
    with zero mean and variance \( \sigma^2 \). 
  \end{remark}

  \section{General linear models.} We are often interested in finding the
  best model function in a specific function space.  Assuming this space has a
  basis \( \basis = \set{\varphi_i}_{i=1}^M\), we may write our model function
  in terms of the basis functions and the model parameters as
  \begin{equation}
    \label{eq:gen_lin_mod}
    f(\x, \beta) \coloneqq \sum_{i = 1}^M \beta_i\varphi_i(\x).
  \end{equation}
  With this formulation, we can represent the approximation \( \hat{\y} \) to
  \( \y \) as a matrix product 
  \begin{equation}
    \hat{\y} = \X\beta
  \end{equation}
  where \( \X \) is the \emph{design matrix} defined by components \(\X_{ij} =
  \varphi_j(\x_i)\), and \( \beta = [\beta_0, \ldots, \beta_M] \) is the model
  parameters. With this notation in mind, we can also write the cost function
  as
  \begin{equation}
    \cost(\data, \beta) = \frac{1}{n} (\y - \hat{\y})^T (\y - \hat{\y}).
  \end{equation}

  \paragraph{Optimizing the parameters of a general linear model.}

  Assume now that our model is of the form given in \cref{eq:gen_lin_mod} and
  that our cost function is the mean squared error defined in \cref{eq:mse}. We
  are interested in finding the parameters \( \beta \) that minimize the cost
  function \( \cost(\data, \beta) \). Since \( \cost(\data, \beta) \) is
  convex, it suffices to differentiate with respect to \( \beta \) and equating
  to zero.

  We have that
  \begin{align}
    \dpd{\cost(\data, \beta)}{\beta} = \X^T(\y - \X\beta), 
  \end{align}
  and equating this to zero yields the familiar \emph{normal equations}
  \begin{equation}
    \X^T \y = \X^T\X\beta.
  \end{equation}
  If the matrix \( \X^T\X \) is invertible, we may obtain the solution by
  direct numerical matrix inversion. In this case, the optimal model parameters
  are found directly by
  \begin{equation}
    \beta = (\X^T\X)^{-1}\X^T\y.
  \end{equation}
  However, these matrices may be ill-conditioned when the number of equations
  are very large, and it is therefore common to apply approximate solvers for
  the inverse, for instance using low-rank approximation based on
  \textsc{SVD}-decomposition, which we will briefly turn to in the following.

  \paragraph{Singular value decomposition.}

  Recall that any matrix \( \mat{A} \in \C^{n, m} \) can be decomposed as
  \begin{equation}
    \mat{A} = \mat{U} \mat{\Sigma}\mat{V}^T,
  \end{equation}
  where \( \mat{U} \) and \( \mat{V} \) comprised of the eigenvectors of \(
  \mat{A}\mat{A}^T \) and \( \mat{A}^T\mat{A} \) respectively. As these
  eigenvectors are orthonormal, it follows that both \( U \) and \( V \) are
  unitary matrices. Furthermore, \( \mat{\Sigma} \) is a matrix 
  \begin{equation}
    \mat{\Sigma} = \begin{bmatrix}
      \Sigma_1 & 0 \\
       0 & 0
    \end{bmatrix},
  \end{equation}
  where \( \Sigma_1 \) is a square diagonal matrix of size \( r \times r \)
  with the non-zero singular values of \( \mat{A} \). The integer \( r \) is
  the \emph{rank} of \( \mat{A} \). As the matrix \( \Sigma \) is mostly
  containing zeros, the information stored in \( \mat{A} \) is attributed to
  only some parts of \( \mat{U} \) and \( \mat{V} \). We can remove the
  redundant parts, and more compactly express \( \mat{A} \) as \( \mat{A} =
  \mat{U}_1 \mat{\Sigma}_1 \mat{V}_1^T \). Here, \( \mat{U}_1 \) is \( m \times
  r \) and \( \mat{V}_1 \) is \( n \times r \).
\end{document}