### 2019-12-12 assignment_two/article/main.tex
Working on assignment two

\documentclass[a4paper, oneside, article]{memoir}

\usepackage{minted} 
\usepackage[T1]{fontenc} 
\usemintedstyle{algol}
\usepackage[]{microtype}
\usepackage[]{eulervm}
\usepackage[]{palatino} 
\usepackage[]{amsmath, amssymb} 
\usepackage[]{graphicx} 
\usepackage[]{svg} 
\usepackage[]{bm} 
\usepackage[]{commath} 
\usepackage[]{geometry} 
\usepackage[]{hyperref} 
\usepackage[]{cleveref}

\setlength{\parskip}{1em}
\setlength{\parindent}{0pt}

\newcommand{\OLS}{\mathrm{OLS}}
\newcommand{\Ridge}{\mathrm{ridge}}
\newcommand{\Lasso}{\mathrm{lasso}}
\newcommand{\X}{\bm{X}}
\newcommand{\y}{\bm{y}}
\newcommand{\yhat}{\hat{\bm{y}}}
\newcommand{\transpose}[1]{#1^{\top}}
\newcommand{\mat}[1]{\bm{#1}}
\newcommand{\cost}{\mathcal{C}}

\title{\textsc{Regression and Classification using Feed Forward Neural Networks}}
\author{Ivar Stangeby}
\begin{document}
	\maketitle

	\begin{abstract}

	\end{abstract}
	\chapter{Introduction}

	In this project we make a comparative study on feed forward neural
	networks in a regression and classification setting. In particular, we
	compare feed forward neural networks in this context with the standard
	methods of ordinary least squares for linear regression, and logistic
	regression for classification.

	\chapter{Background}

	\section{Neural Networks}

	\subsection{Construction}
	Neural networks has gained a widespread use in the approximation of
	functional relationships. In it's simplest form, a neural network
	consists of a set of \emph{neurons}, being entities that take as input
	a set number of parameters \( x_1, \ldots, x_n \), and computes a
	linear combination of these (often with added bias):
	\begin{equation}
		z = \sum_{i = 0}^n w_ix_i + b.
	\end{equation}
	The
	neuron then decides on whether to ``fire'' or not, by evaluating a
	so-called scalar activation function, \( a = g(z) \). In the architecture
	discussed herein, namely \emph{fully connected} or \emph{dense} neural
	networks, each neuron takes into account all possible parameters.

	By combining a set number of neurons we obtain what is called a
	\emph{layer}. A neural network typically consists of three types of
	layers: an \emph{input layer}, a set number of \emph{hidden layers},
	and finally an \emph{output layer}.

	\begin{figure}[htbp]
		\centering
		\includegraphics[width=0.8\textwidth]{imgs/fcc_nn.pdf}
		\caption{A simple feed forward neural network. This specific
		network has an input layer of four nodes. Two hidden layers
	consisting of six nodes each, and finally, an output layer of size
two.}%
\label{fig:fc_nn}
	\end{figure}

	For a network consisting of \( L \) layers, with each layer having \(
	n^{[\ell]} \) nodes and with activation function \( g\), the
	activations in each layer are given as 
	\begin{equation}
		a^{[\ell]}_k = g\left(\sum_{j = 1}^{n^{[\ell]}} w_{jk}^{[\ell]} a^{[\ell - 1]}_j  + b^{[\ell]}_k \right)
	\end{equation}
	for \( k = 1, \ldots, n^{[\ell]} \) and \( \ell = 1, \ldots, L \). 
	A matrix formulation reads:
	\begin{equation}
	\mat{a}^{[\ell]} = g\left(\mat{W}^{[\ell]} \mat{a}^{[\ell - 1]} + \mat{b}^{[\ell]}\right).
	\end{equation}
	We use the convention that the network input \( \mat{x} = (x_1, \ldots,
	x_{n^{[0]}}) = \mat{a}^{[0]} \) and the network output \( \yhat =
	\mat{a}^{[L]} \).

	\subsection{Optimization}

	We initialize such a network by specifying the initial weights and
	biases for each layer. That is, the matrices \( \mat{W}^{[\ell]} \) and
	\( \mat{b}^{[\ell]} \) are (most commonly) drawn from some random
	distribution. It is our goal to iteratively optimize these weights and
	biases in order for the network to produce desired output. We therefore
	choose a \emph{cost} function  \( \cost \) which in some way
	incorporates our problem at hand. For example, in regression problems,
	our cost function is typically the mean squared error (MSE) between our
	target-data and our prediction, while in classification, we often use
	cross entropy loss.

	We optimize the network by computing the gradient of the cost function
	\( \cost \) with respect to weights and bias' and perform gradient
	descent. That is
	\begin{align}
		\begin{split}
			\mat{W}^{[\ell]} \leftarrow &\mat{W}^{[\ell]} - \lambda \dpd{\cost}{{\mat{W}^{[\ell]}}}, \\ 
			\mat{b}^{[\ell]} \leftarrow &\mat{b}^{[\ell]}\;\, - \lambda \dpd{\cost}{{\mat{b}^{[\ell]}}}
		\end{split}
	\end{align}
	Here \( \lambda > 0 \) is the \emph{learning rate}.

	\subsection{Neural Network Architecture}

	In this paper we study two use-cases of feed forward neural networks,
	namely regression, and classification. The architectures are tailored
	to the data-sets at hand.

	\paragraph{Credit Card Data}

	The credit-card data-set consists of 33 numerical features describing a
	users payment history.  The task at hand is then to predict whether the
	user is prone to defaulting or not. This is therefore a binary
	classification problem. Our network therefore needs to have an input
	layer of size 33, and an output layer of size two (one probability
	value for each class). We---rather arbitrarily---decide on three hidden
	layers of size 20. We use the \emph{rectified linear unit} as our
	activation function for the hidden-layers, and (since we are predicting
	probabilities) we use the \emph{softmax} function as our
	output-activation. These are defined as follows:
	\begin{align}
		\mathrm{RELU}(\mat{x}) &= \max\{\mat{0}, \mat{x}\}, \\
		\mathrm{Softmax}_k(\mat{x}) &= \frac{\exp{x_k}}{\sum_{i=1}^n \exp{x_i}}.
	\end{align}
	Note that the softmax takes as input a vector, and spits out a scalar,
	and that the RELU takes the element-wise maximum.
	Our cost function \( \cost \) is chosen to be \emph{binary cross
	entropy} (or \emph{log-loss}) defined as
	\begin{equation}
		\cost(\mat{y}, \yhat) = -(\mat{y} \log(\yhat) + (1 - \mat{y}) \log(1 - \yhat)).
	\end{equation}

	\paragraph{Regression Data}

	For our regression task, we create a synthetic data-set by noisy
	sampling of points from the Franke test-function, and our goal is to
	create a neural network that approximates the Franke function. Thus,
	our network should take as input a set of two values, and return a
	scalar. Again, we use a rectified linear unit as our activation
	function for the hidden layers, however, since we want an unrestrained
	output range, we do not add the softmax activation to the final layer,
	as this would squeeze the network output between zero and one.
	Our cost function is the mean squared error (MSE):
	\begin{equation}
		\cost(\mat{y}, \yhat) = \frac{1}{n} (\y - \yhat)^T (\y - \yhat).
	\end{equation}

	\chapter{Implementation}

	While implementing backpropagation is a great exercise in programming,
	due to time-constraints, we have had to resort to using machine
	learning frameworks in the implementation.

	For defining our neural networks we use the Pytorch-library, and for
	standard methods like linear regression and logistic regression, we
	resort to the scikit-learn toolkit. In addition to this, we use the
	Skorch-library for wrapping the Pytorch modules in a scikit-learn
	compatible API.

	\subsection{Data Wrangling}

	We subclass the PyTorch dataset-class for loading the credit card data
	and the sampled Franke data as shown in \cref{lst:datasets}. This
	enables us to write lean training loops as the \texttt{DataLoader}
	class works as a batched iterator.

	\subsection{Neural Network}

	We implement a generic neural network that will suffice for both the
	credit card data (the classification task) and the Franke sampled data
	(the regression task). By subclassing \texttt{torch.nn.Module}, we only
	need to register the different layers, and implement the
	\texttt{forward}-method. The auto-grad framework then takes care of the
	back-propagation of gradients for us. We allow for the user to specify
	the number of input features, the number of output features, the number
	of hidden layers and the the number of hidden units (which we take as
	constant across the hidden layers). Furthermore, we also with to be
	able to specify the final activation function, as this needs to be
	specially tailored (as much else) to the task at hand. The generic architecture is layed out in
	\cref{lst:neural_network}.

	\begin{listing}
	\begin{minted}[frame=lines,tabsize=3]{python}
class CreditCardData(torch.nn.utils.data.Dataset):
	def __init__(self, *args, **kwargs):
		super().__init__()

		self.X, self.y = self.fetch_data(*args, **kwargs)
		self.num_features, self.num_items = self.X.shape

	def __getitem__(self, index):
		return self.X[index], self.y[index]

	def __len__(self):
		return self.num_items

		super().__init__()	
	def fetch_data(self):
		# Handles the loading and preprocessing of data
		...

class FrankeData(torch.nn.utils.data.Dataset):
	...
	\end{minted}
	\caption{Interfacing the data-sets with a PyTorch API, which facilitates easy iteration as shown in \cref{lst:training_loop}}
	\label{lst:datasets}
	\end{listing}

	\begin{listing}
	\begin{minted}[frame=lines,tabsize=3]{python}
credit_card_data = CreditCardData('path_to_dataset')
credit_card_loader = DataLoader(credit_card_data, batch_size=64)

# training loop
for (X_batch, y_batch) in credit_card_loader:
	...
	\end{minted}
	\caption{A simple training loop using PyTorch data-loaders.}
	\label{lst:training_loop}
	\end{listing}

	\begin{listing}
	\begin{minted}[frame=lines,tabsize=3]{python}
class GenericNN(torch.nn.Module):

	def __init__(
		self, num_in_features, num_out_features,
		num_hidden_layers=3,
		num_hidden_features=20,
		activation=torch.nn.Relu,
		final_activation=torch.nn.Identity
	):	
		super().__init__()
		self.input_layer = torch.nn.Linear(num_in_features, 
			num_hidden_features)
		self.hidden_layers = torch.nn.ModuleList(
		[
			nn.Linear(num_hidden_features, num_hidden_features)
			for i in range(num_hidden_layers) 
		])
		self.output_layer = torch.nn.Linear(num_hidden_features,
			num_out_features)
		self.g = activation
		self.f = final_activation

	def forward(self, x):
		x = self.g(self.input_layer(x))
		for layer in self.hidden_layers:
			x = self.g(layer(x))
		x = self.f(self.output_layer(x))

		return x
	\end{minted}	
	\caption{A generic dense neural network realized in the PyTorch
	framework. By implementing the forward method, gradients are kept track
	of automatically, and hence no consideration is to be made for the backward
	propagation.}
	\label{lst:neural_network}
	\end{listing}

	\subsection{Classical Methods}

	For the classical methods of (regularized) ordinary least squares and
	logistic regression, we simply use the scikit-learn toolkit. These
	classifiers can be trained by calling \texttt{fit}-method. By utilizing
	cross validation methods, we optimize the hyper-parameters. This will
	be discussed further in the results-section. As an example, for the credit-card dataset, we
	train the classifiers shown in \cref{lst:classifier_credit_card}, performing a grid-search for the best regularization strength:
	\begin{listing}
	\begin{minted}[frame=lines, tabsize=3]{python}
l1_params = [{'penalty':['l1'],'C':np.logspace(-4, 4, 40),\
		'solver':['liblinear']}]
l2_params = [{'penalty':['l2'],'C':np.logspace(-4, 4, 40),\
		'solver':['liblinear']}]

clf_l1 = GridSearchCV(LogisticRegression(), l1_params, cv=5,\
	verbose=True, n_jobs=-1)
clf_l2 = GridSearchCV(LogisticRegression(), l2_params, cv=5,\
	verbose=True, n_jobs=-1)

clf_l1.fit(X, y)
clf_l2.fit(X, y)
	\end{minted}
	\caption{Setting up and training the logistic regressor. We use both
	\(L^1\) and \( L^2\)-regularization, corresponding to Lasso and
Ridge-regression, respectively. We find the optimal regularization-strength by
5-fold cross validation over the data-set. The fits are parallelized over all
available CPU-cores, to speed up the process.}
	\label{lst:classifier_credit_card}
	\end{listing}

	\section{Numerical Results}

	In this section, we present the findings from our numerical experiments.

	\subsection{Credit card data}

	We train our sklearn-classifiers as in
	\cref{lst:classifier_credit_card}. We do a hyper-parameter search for
	the best regularization-strength for both \( L^1 \) and \( L^2 \)
	regularization. We note that \emph{regularization strength} in the
	world of scikit-learn is the inverse of what might be the convention.
	Thus, a regularizatin strength of \( \lambda \) corresponds to the
	penalization coefficient \(  1 / \lambda \). With this in mind, we
	display the results in \cref{fig:credit_cards_classical}. Note how, for
	small values of \( \lambda \), the regularization effectively sets all
	weights to zero. The fact that the network still predicts correctly
	about 78.6\% of the time is due to the fact that the network now only
	predicts zeros, and the data-set consists of about 78.6\% zeros. In the
	other limit, as \( \lambda \) increases, the regularization term is
	effectively set to zero, and both methods converge to non-penalized
	logistic regression. Incidentally, it is also this non-penalized method
	that performs the best with a classification accuracy of \( 81.6 \% \).

	The neural network classifier is trained using the
	\texttt{skorch}-framework, which acts as an interface to scikit-learn
	for PyTorch. Using this allows for  simple hyper-parameter optimization
	as in scikit-learn, and one may test for instance network depth, width,
	and regularization strength. Unfortunately, due to time constraints, we
	were only able to perform a learning-rate optimization. However, the
	implementation in \cref{lst:classification_nn} allows for other sets of
	hyper-parameters to be tested.

	\begin{figure}[htpb]
		\centering
		\includegraphics[width=0.8\linewidth]{imgs/credit_card_classification.pdf}
		\caption{The results of the \( L^1 \) and \( L^2 \) regularized logistic regression. 
			The two regularization types seems to converge to
			exactly the same classification accuracy as
			non-penalized logistic regression, with this value being 81.6\%. As \( \lambda \)
			increases, the penalization term is set to zero, hence
			converges to non-penalized regression. Why this
			coincides with the best fit is not clear to me.
		}%
		\label{fig:credit_cards_classical}
	\end{figure}

	\begin{listing}
	\begin{minted}[frame=lines, tabsize=3]{python}
net = RegularizedNeuralNetClassifier(
    GenericNN,
    module__input_size=credit_card_data.num_features,
    module__num_hidden_layers=3,
    module__output_size=2,
    criterion=torch.nn.CrossEntropyLoss,
    train_split=CVSplit(cv=5, random_state=42),
    optimizer=torch.optim.SGD,
    max_epochs=5,
    optimizer__momentum=0.9
)

params = {
    'lr': np.logspace(-4, 0, 30),
    # these hyper-params can be trained for as well
    # 'module__num_hidden_layers' : range(0, 5),
    # 'module__hidden_layer_size' : [5, 10, 15, 20]
    # 'module__regularization = ['l2', 'none', 'l1'] 
    # 'module__alpha = np.logspace(-10, 10, 40) # regularization str
}
clf = GridSearchCV(
    net,
    params,
    n_jobs=12,
    cv=5
)
clf.fit(credit_card_data.X.numpy(), credit_card_data.y.numpy())
	\end{minted}	
	\caption{By wrapping the GenericNN in a RegularizedNeuralNetClassifier
	(see \cref{lst:rnnc}) we may train the PyTorch model using the
	GridSearchCV-functionality from scikit-learn.}
	\label{lst:classification_nn}
	\end{listing}

	The results are shown in \cref{fig:credit_card_nn}, and as we see, the
	neural network, even with its basic architecture, outperforms the
	classical methods by a fair bit. The optimal learning rate yields a
	classification accuracy of 82.54\%.

	\begin{figure}[htpb]
		\centering
		\includegraphics[width=0.8\linewidth]{imgs/classification_credit_cards_nn.pdf}
		\caption{The classification accuracy as a function of the
			learning rate for the generic neural network. At the
			opimal learning rate, the network achieves a prediction
			accuracy of approximately 82.54\%, which is not
			insignificantly better than all the classical methods.
			The filled blue regions represent the standard
			deviations as reported by \texttt{GridSearchCV}. Note
			how the network at performs it's worst by guessing only
			zeros, similarily to the overpenalized logistic
		regression. The network is trained over 20 epochs, using SGD
	with momentum and 5-fold cross validation. }%
		\label{fig:credit_card_nn}
	\end{figure}
\appendix

\chapter{Regularized neural network}
	\begin{listing}
	\begin{minted}[frame=lines, tabsize=3]{python}
class RegularizedNeuralNetClassifier(NeuralNetClassifier):

    def __init__(self, module, alpha=1, regularizer='none', 
    		criterion=torch.nn.NLLLoss,
                train_split=CVSplit(5, stratified=True),
                classes=None,
                *args,
                **kwargs):
        self.alpha = alpha
        self.regularizer = regularizer

        if 'regularizer' in kwargs.keys():
            kwargs.pop('regularizer')
        if 'alpha' in kwargs.keys():
            kwargs.pop('alpha')

        super().__init__(module, *args, criterion=criterion, \
		train_split=train_split, classes=classes, **kwargs)

    def get_loss(self, y_pred, y_true, *args, **kwargs):
        loss = super().get_loss(y_pred, y_true, *args, **kwargs)

        if self.regularizer == 'none':
            pass
        elif self.regularizer == 'l1':
            loss += self.alpha * sum([w.abs().sum() 
	    					for w in self.module_.parameters()])
        elif self.regularizer == 'l2':
            loss += self.alpha * sum([(w.abs() ** 2).sum() 
	    					for w in self.module_.parameters()])
        return loss

	\end{minted}	
	\caption{We implemented a regularized neural network classifier in
	order to be able to compare the network with both \( L^2 \) and \( L^1
	\) regularization applied to the network weights. Unfortunately, we were not
	able to perform the required numerical tests in time, as the parameter space is
	rather large.}
	\label{lst:rnnc}
	\end{listing}

	\subsection{Franke function}
\end{document}