### 2019-12-20 assignment_three/article/main.tex
Working on project 3.

\documentclass[article, a4paper, oneside]{memoir}
\linespread{1.25}

\setlength{\parskip}{1em}
\setlength{\parindent}{0em}
\usepackage[]{eulervm} 
\usepackage[]{palatino} 
\usepackage[]{commath} 
\usepackage[]{microtype} 
\usepackage[]{geometry} 
\usepackage[]{minted} 
\usemintedstyle{algol}

\usepackage[]{amsmath, amssymb, amsthm, mathtools} 
\usepackage[backend=biber, bibencoding=utf8, style=ieee]{biblatex} 
\addbibresource{../../FYS-STK4155.bib}

\usepackage[]{hyperref} 
\usepackage[noabbrev, capitalize]{cleveref}

\newcommand{\cost}{\mathcal{C}}

\title{\textsc{Solving the heat-equation using neural networks}}
\author{Ivar Stangeby}

\begin{document}
	\maketitle

	\chapter{Introduction}

	Partial differential equations have widespread use in the sciences, in
	fields ranging from structural engineering to economics. The solution
	of such equations in an accurate and precise way is of paramount
	importance.

	Several approaches to the solution of such equations exists, with the
	more notable being \emph{finite difference} and \emph{finite
	element}-methods. These two methods approach the problem at hand from
	two different angles. In the finite difference schemes, the
	differential operator is discretized, typically using schemes arising
	from Taylor approximations to function. Finite element methods on the
	other hand, instead of discretizing the operators, discretize the
	function space for which the solution is sought.

	In recent years, the use of artificial intelligence in the solution of
	partial differential equations has started to gain some traction, see
	for instance the DeepXDE-network \cite{luDeepXDEDeepLearning2019} and
	the Deep Galerkin Method (DGM)
	\cite{al-aradiSolvingNonlinearHighDimensional}.

	The neural network approach to solving PDEs is meshless, that is, it
	does not rely on any underlying discretization of the domain as opposed
	to finite difference and finite element methods. Furthermore, the
	boundary conditions of the problem are encoded directly in the
	cost-function used to optimize the neural network. In finite elements
	the solution space is constructed in such a way as to automatically
	satisfy the boundary conditions, and in finite differences the boundary
	conditions must be imposed explicitly.

	\chapter{Theory}

	\section{The heat equation}

	We wish to solve the one-dimensional heat-equation. The problem reads
	as follows. Find \( u \) such that
	\begin{equation}
		\label{eq:heat_eq}
		\pd[2]{u(x, t)}{x} = \pd{u(x, t)}{t}
	\end{equation}
	subject to the boundary conditions \( u(0, t) = u(1, t) = 0 \) and the
	initial condition
	\begin{equation}
		u(x, 0) = \sin(\pi x).
	\end{equation}
	We interpret the problem as that of finding the temperature gradient in
	a rod of fixed length \( L = 1\).

	\subsection{An analytical solution}

	In order to assess a-posteriori model performance, we need to know the
	exact solution to our specific instance of the heat equation.

	Following \cite{hancockLinearPartialDifferential} we derive an
	analytical solution.  We assume separation of variables and see whether
	this leads us to anything conclusive. We write
	\begin{equation}
		u(x, t) = X(x)T(t)
	\end{equation}
	where \( X \) carries the spatial dependence and \( T \) carries the
	temporal dependence of the problem.  By taking the required partial
	derivatives we obtain \(X''(x)T(t) = X(x)T'(t)\) from which we can
	deduce that the ratios
	\begin{equation}
		\label{eq:ratio}
		\frac{X''(x)}{X(x)} = \frac{T'(t)}{T(t)} = -\lambda
	\end{equation}
	are constant. From the boundary conditions, we have that
	\begin{align}
		u(0, t) = X(0)T(t) = 0, \\
		u(1, t) = X(1)T(t) = 0.
	\end{align}
	By examination, we note that if \(T(t) = 0\) we obtain the trivial
	solution \( u = 0\), which does not satisfy the boundary conditions.
	Hence \( T \neq 0 \). Thus to satisfy the boundary conditions we must
	have \( X(0) = X(1) = 0 \).

	The ratios in \cref{eq:ratio} gives rise to two ODEs, one in time and
	one in space:
	\begin{align}
		X''(x) + \lambda X(x) &= 0, \label{eq:spatialODE} \\
		T'(t) + \lambda T(t) &= 0 \label{eq:temp_ODE}.
	\end{align}
	We have obtained boundary conditions for the spatial ODE, hence we
	start by solving that.

	As the constant \( \lambda \) is unknown, we have to consider three
	cases. The cases \( \lambda < 0 \) and \( \lambda = 0 \) can be shown
	to lead to \( u = 0\) which we discard. Thus we assume \( \lambda >
	0\).  The solution to \cref{eq:spatialODE} is then
	\begin{equation}
		X(x) = A\cos(\sqrt{\lambda}x) + B\sin(\sqrt{\lambda}x).
	\end{equation}
	By imposing the boundary conditions we obtain \( A = 0 \) and \(
	B\sin(\sqrt{\lambda}) = 0 \). Again, if we allow \( B = 0\), we obtain
	\( u = 0\), so we discard this possibility. Hence \( \sin(\sqrt{\lambda}) = 0
	\) from which we can deduce that
	\begin{equation}
		\lambda = \pi^2 n^2 
	\end{equation}
	for \( n \in \mathbb{N}^+\). This tells us that we have an infinite
	number of solutions, one for each \( n \):
	\begin{equation}
		\label{eq:solution_space}
		X_n(x) = B_n\sin(\pi n x)
	\end{equation}
	for some unknown constant \( B_n \).

	We now solve for the spatial component \( T \). For each possible
	choice of \( \lambda \) we obtain
	\begin{equation}
		T_n'(t) +\pi^2 n^2T_n(t) = 0, 
	\end{equation}
	which has solutions
	\begin{equation}
		\label{eq:solution_temp}
		T_n(t) = C_n e^{-n^2\pi^2t}.
	\end{equation}
	Combining \cref{eq:solution_space,eq:solution_temp} we get the family of solutions
	\begin{equation}
		u_n(x, t) = B_n C_n \sin(n\pi x)e^{-n^2\pi^2t}.
	\end{equation}
	In principle, we would have to consider the fourier coefficients of the
	infinite sum of these solutions, however, by examining our initial
	conditions, we see that \( u_1 \) is the desired solution, with \( B_1
	C_1 = 1 \).

	Thus, our closed form solution to the one-dimensional heat equation
	with our prescribed boundary and initial conditions is:
	\begin{equation}
		u(x, t) = \sin(\pi x) e^{-\pi^2 t}.
	\end{equation}

	\section{Finite differences}

	We wish to solve the heat equation over the domain \( \Omega = [0, L]
	\times [0, T]  \). Discretizing using \( N \) grid points in the
	spatial direction and \( M \) grid points in the temporal direction, we
	obtain step sizes \( \Delta x = L / (N - 1) \) and \( \Delta t = T / (M
	- 1) \). Of the myriad of different finite difference schemes, we
	choose the \emph{forward in time---centered in space} (FTCS).

	\subsection{Forward in time}
	We start by discretizing the time-differential in \cref{eq:heat_eq} by
	taking a first-order Taylor approximation of \( u \). We have
	\begin{equation}
		u(x, t + \Delta t) = \sum_{n = 0}^\infty \frac{\Delta t^n}{n!} \pd[n]{u(x, t)}{t} = u(x, t) + \Delta t \pd{u(x, t)}{t} + \mathcal{O}(\Delta t).
	\end{equation}
	We discard higher order terms and rearrange, yielding
	\begin{equation}
		\label{eq:forward_time}
		\pd{u(x, t)}{t} \approx \frac{u(x, t+\Delta t) - u(x, t)}{\Delta t}, 
	\end{equation}
	with a truncation error that goes as \( \mathcal{O}(\Delta t) \).

	\subsection{Centered in space}

	The centered difference scheme involves two second-order
	Taylor-aproximations in space:
	\begin{align}
		u(x + \Delta x, t) &= u(x, t) + \Delta x \dpd{u(x, t)}{x} +
		\frac{\Delta x^2}{2} \dpd[2]{u(x, t)}{x} + \mathcal{O}(\Delta x^2) \\
		u(x - \Delta x, t) &= u(x, t) - \Delta x \dpd{u(x, t)}{x} + \frac{\Delta x^2}{2} \dpd[2]{u(x, t)}{x} + \mathcal{O}(\Delta x^2).
	\end{align}
	Truncating, adding the equations to elimiate the first order
	derivatives, and solving for the second derivative yields:
	\begin{equation}
		\label{eq:centered_space}
		\dpd[2]{u(x, t)}{x} \approx \frac{u(x + \Delta x, t) - 2 u(x,
		t) + u(x - \Delta x, t)}{\Delta x^2}
	\end{equation}
	with a truncation order of \( \mathcal{O}(\Delta x^2) \).

	\subsection{The full scheme}

	Plugging \cref{eq:forward_time,eq:centered_space} into the heat
	equation given in \cref{eq:heat_eq} yields
	\begin{equation}
		\frac{u(x, t + \Delta t) - u(x, t)}{\Delta t} = \frac{u(x +
		\Delta x, t) - 2u(x, t) + u(x - \Delta x, t)}{\Delta x^2}.
	\end{equation}
	We are interesting in solving the equation forwards in time, so we
	solve for the term \( u(x, t + \Delta t) \), yielding:
	\begin{equation}
		u(x, t + \Delta t) = \left( u(x + \Delta x, t) + u(x - \Delta x, t) \right) \frac{\Delta t}{\Delta x^2} + \left(1 - 2 \frac{\Delta t}{\Delta x^2}\right)u(x, t).
	\end{equation}
	Note that the solution at the next time step requires three spatial
	solutions at the previous time-step. Hence, we are reliant on the
	boundary conditions to keep the algorithm running.

	While this scheme is easy to implement, it suffers from stability
	issues, which can be combated by tweaking the ratio \( \Delta t /
	\Delta x^2\). By ensuring that 
	\begin{equation}
		\Delta t \leq \frac{\Delta x^2}{2}, 
	\end{equation}
	the method behaves nicely.

	\section{Neural network in the context of PDEs}

	Consider a neural network \( f:\Omega \to \mathbb{R}\) parameterized by \(\theta\).  The
	question is now, how do we embed the partial differential equation in
	the neural network? The answer lies in how we engineer our loss
	function.

	\subsection{Criterion}

	From \cref{eq:heat_eq} we can see that we wish to minimize
	the following expression:
	\begin{equation}
		\label{eq:min_residual}
		\min_{\theta} \left[\dpd[2]{f(x, t; \theta)}{x} - \dpd{f(x, t; \theta)}{t}\right]
	\end{equation}
	for \( x \in [0, L] \) and \( t > 0\). Thus, we can simply use the mean
	squared error of the residuals in the right hand side of
	\cref{eq:min_residual}. We therefore define our cost function \( \cost
	\) as the mean squared residual error over all (assume \( n \))
	data-points  in \(\Omega = [0, L] \times [0, T]\).
	\begin{equation}
		\cost(f) = \frac{1}{n} \sum_{i=1}^n \left(  \dpd[2]{f(x_i, t_i)}{x}
		- \dpd{f(x_i, t_i)}{t}\right)^2
	\end{equation}

	\subsection{Embedding boundary conditions}

	Currently, we are not imposing any boundary conditions. Thus, by
	naively running the network as is, we may find any function whose
	second partial derivative in space is equal to its partial derivative
	in time. There are several approaches to solving this problem. In
	\cite{luDeepXDEDeepLearning2019} they train the network according to a
	multi-task loss following the same trick as in \cref{eq:min_residual}
	but for data-points lying on the boundary.

	We instead opt for the method of constructing a \emph{trial function}
	that acts as a mediator between the neural network and the
	loss-function which ensures that the network learns to satisfy the
	boundary conditions.

	Recall from our boundary and initial conditions that we require \( u(0,
	t) = u(1, t) = 0 \) and \( u(x, 0) = \sin(\pi x) \). The function \( g
	\) defined by
	\begin{equation}
		\label{eq:trial}
		g(x, t; \theta) = (1 - t) \sin(\pi x) + x(1-x)t f(x, t; \theta),
	\end{equation}
	satisfies these conditions. It is therefore this function we pass into
	our cost function \( \cost \).

	\subsection{Neural network architecture}

	As our architecture we wanted to reuse the \textsc{GenericNN}
	implemented in \cite{stangebyFYSSTK4155Project2019a}, which is a
	feed-forward neural network with \textsc{Relu}-activations at each
	hidden layer. However, after initial testing, using \textsc{Relu}
	yielding wildly oscilating loss, hence we chose the more stable
	\textsc{Sigmoid}-activation. This led to slower learning rates, due to
	the saturation of gradients, but allowed the network to get a tighter
	approximation.

	\section{Performance metrics}

	In order to evaluate our numerical approximations, whether they are
	computed using finite-difference or they are trained neural networks,
	we have several metrics we can consider. Calling our numerical
	approximation to \( u \) for \( \tilde{u} \), we define our errors as
	\begin{equation}
		e_p = \| u - \tilde{u} \|_p,
	\end{equation}
	where \( \| \cdot \|_p \) denotes the \(p\)-norm for vectors.

	\chapter{Implementation}

	The finite difference solver is implemented using a naive for-loop
	approach in \texttt{numpy} and can be seen in \cref{lst:ftcs}. In order
	to satisfy the stability criterion, we choose our desired \( \Delta x
	\) and compute a corresponding \( \Delta t\) by
	\begin{equation}
		\Delta t = \Delta x^2 / 2.
	\end{equation}

	The neural network is implemented in \texttt{tensorflow}, which takes
	care of the backpropagation of gradients through the network using the
	\texttt{AutoGrad}-package. Ideally, we wanted to implement this in
	\texttt{PyTorch}, however, taking element-wise gradients of the network
	output with the respect to the network input, as is required to
	evaluate \cref{eq:trial}, turned out to be hard to do in
	\texttt{PyTorch}. Due to my lacking experience in \texttt{Tensorflow},
	the implementation is a modified version of the one used to solve the
	wave equation in \cite{heinDataAnalysisMachine}. This was implemented
	in \texttt{Tensorflow V1}, thus we had to add the compatibility
	modifications shown in \cref{lst:compatibility} in order for this to
	run using the newer \texttt{Tensorflow V2} API.

	\begin{listing}
		\centering	
		\begin{minted}{python}
import tensorflow.compat.v1 as tf
tf.disable_eager_execution()		
		\end{minted}
		\caption{Compatibility modification needed to use the
		deprecated functionality of \texttt{Tensorflow} V1 in the V2
	API.}
		\label{lst:compatibility}
	\end{listing}

	\begin{listing}
		\centering
		\begin{minted}{python}
def ftcs(space_resolution, time_resolution, space_min_max=[0, 1],
		time_max=1, boundary_conditions=[0, 0], initial_condition=None):
    """
    Solves the 1D-heat equation using a forward-in-time 
    centered-in-space discretization scheme.

    :param space_resolution: number of points in spatial direction
    :param time_resolution: number of points in temporal direction
    :param space_min_max: the spatial boundary values
    :param time_max: the time to run the simulation for
    :param boundary_conditions: the boundary conditions at the space_min_max-values
    :param initial_conditions: the initial conditions at time = 0, callable.
    """

    if initial_condition is None:
        initial_condition: lambda x: 0

    dt = time_max / time_resolution
    u0, dx = np.linspace(*space_min_max, num=space_resolution, retstep=True)
    u = np.zeros((time_resolution, space_resolution))
    u[:, [0, -1]] = boundary_conditions
    u[0] = initial_condition(u0)

    F = dt / dx ** 2
    for step in tqdm.trange(time_resolution - 1):
        for i in range(1, space_resolution - 1):
            u[step + 1, i] = u[step, i] +\ 
	    	F * (u[step, i - 1] - 2 * u[step, i] + u[step, i + 1])

    return u

@np.vectorize
def initial_condition(x):
    return np.sin(np.pi * x)
		\end{minted}
		\caption{A \texttt{numpy}-based implementation of the
			\emph{forward in time---centered in space} scheme. The
			initial conditions are passed in as a callable, and the
			function returns the solution at all time-steps in form
			of a \texttt{time\_resolution} \(\times\)
		\texttt{space\_resolution} matrix.}
		\label{lst:ftcs}
	\end{listing}

	\chapter{Numerical experiments}

\clearpage
\printbibliography
\end{document}